---
title: "practical_exercise_2, Methods 3, 2021, autumn semester"
author: 'Jessica'
date: "22/9-21"
output: pdf_document
---

<style type="text/css">
  body{
  font-size: 14pt;
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 1: Using mixed effects modelling to model hierarchical data
In this assignment we will be investigating the _politeness_ dataset of Winter and Grawunder (2012) and apply basic methods of multilevel modelling. 

## Dataset
The dataset has been shared on GitHub, so make sure that the csv-file is on your current path. Otherwise you can supply the full path.

```{r}
politeness <- read.csv('politeness.csv') ## read in data
library(dplyr)
library(tidyverse)
library(lme4)
library(MuMIn)
```

# Exercises and objectives
The objectives of the exercises of this assignment are:  
1) Learning to recognize hierarchical structures within datasets and describing them  
2) Creating simple multilevel models and assessing their fitness  
3) Write up a report about the findings of the study  

REMEMBER: In your report, make sure to include code that can reproduce the answers requested in the exercises below  
REMEMBER: This assignment will be part of your final portfolio

## Exercise 1 - describing the dataset and making some initial plots

1) Describe the dataset, such that someone who happened upon this dataset could understand the variables and what they contain

-> Subject = participant
-> Gender = gender
-> Scenario = condition (e.g. "apologising for being late", "asking a professor for an extension on an assignment" etc)
-> Attitude = informal or formal (polite) conditions
-> Total_duration = of sentence/saying, in seconds
-> f0mn = mean of the f0, which is something to do with frequency (pitch??)
-> hiss_count = audible and nasal hissing/air-sucking in between talking
We use the glimpse() function from the dplyr package to see what the dataset contains and what the variables are

    i. Also consider whether any of the variables in _politeness_ should be encoded as factors or have the factor encoding removed. Hint: ```?factor```  

```{r}
glimpse(politeness)

#Mutating some of the variables into factors, for later analysis
politeness <- politeness %>% 
mutate(
  gender = as.factor(gender),
  subject = as.factor(subject),
  attitude = as.factor(attitude)
)
```
2) Create a new data frame that just contains the subject _F1_ and run two linear models; one that expresses _f0mn_ as dependent on _scenario_ as an integer; and one that expresses _f0mn_ as dependent on _scenario_ encoded as a factor  
    i. Include the model matrices, $X$ from the General Linear Model, for these two models in your report and describe the different interpretations of _scenario_ that these entail
    ii. Which coding of _scenario_, as a factor or not, is more fitting?
    
```{r}
#Creating a new dataframe with only F1
f1df <- politeness %>% 
  filter(subject == 'F1')

#Model with scenario as an integer
model1 <- lm(f0mn ~ scenario, data = f1df)

#Changing scenario to a factor
f1df <- f1df %>% 
  mutate(
    scenariofactor = as.factor(scenario)
  )

#Model with scenario as a factor
model2 <- lm(f0mn ~ scenariofactor, data = f1df)

model1
model2
#We should keep scenario as a factor, because it doesn't make sense to just look at ONE slope for all scenarios, as they vary

X1 <- model.matrix(model1)
X2 <- model.matrix(model2)
X1
X2

#Again, X2 gives more information
#Scenario should be as a factor, as it makes more sense
```

3) Make a plot that includes a subplot for each subject that has _scenario_ on the x-axis and _f0mn_ on the y-axis and where points are colour coded according to _attitude_
    i. Describe the differences between subjects
```{r}
ggplot(politeness, aes(scenario, f0mn, colour = attitude)) +
  geom_point() +
  facet_wrap(~ subject)

#Males are much lower than females
#Some are not affected by attitude
#Some are affected more by the difference in scenarios
#For the hypotheses that Koreans lower their voices in formal conditions to hold, the red dots should be above the blue, which is hard to conclude anything from in the graphs, but might be true
```


## Exercise 2  - comparison of models

For this part, make sure to have `lme4` installed.  
You can install it using `install.packages("lme4")` and load it using `library(lme4)`  
`lmer` is used for multilevel modelling

```{r, eval=FALSE}
mixed.model <- lmer(formula=..., data=...)
example.formula <- formula(dep.variable ~ first.level.variable + (1 | second.level.variable))
```

1) Build four models and do some comparisons
    i. a single level model that models _f0mn_ as dependent on _gender_
    ii. a two-level model that adds a second level on top of i. where unique intercepts are modelled for each _scenario_
    iii. a two-level model that only has _subject_ as an intercept 
    iv. a two-level model that models intercepts for both _scenario_ and _subject_
    v. which of the models has the lowest residual standard deviation, also compare the Akaike Information Criterion `AIC`?
    vi. which of the second-level effects explains the most variance?
```{r}
m1 <- lm(f0mn ~ gender, data = politeness)
m2 <- lmer(f0mn ~ gender + (1 | scenario), data = politeness)
m3 <- lmer(f0mn ~ gender + (1 | subject), data = politeness)
m4 <- lmer(f0mn ~ gender + (1 | scenario) + (1 | subject), data = politeness)

#m1 has no residuals
#sigma() for finding the residual standard deviation
sigma(m2)
sigma(m3)
sigma(m4)
#m4 has the lowest residual standard deviation

#AIC
AIC1 <- AIC(logLik(m1))
AIC2 <- AIC(logLik(m2))
AIC3 <- AIC(logLik(m3))
AIC4 <- AIC(logLik(m4))

#m4 also has the lowest AIC


#To see which one explains the most variance, we need to find the R^2
r.squaredGLMM(m1)
r.squaredGLMM(m2)
r.squaredGLMM(m3)
r.squaredGLMM(m4)

#Again, m4 has the highest R^2, so it explains the most variance
```

2) Why is our single-level model bad?
    i. create a new data frame that has three variables, _subject_, _gender_ and _f0mn_, where _f0mn_ is the average of all responses of each subject, i.e. averaging across _attitude_ and _scenario_
    ii. build a single-level model that models _f0mn_ as dependent on _gender_ using this new dataset
    iii. make Quantile-Quantile plots, comparing theoretical quantiles to the sample quantiles) using `qqnorm` and `qqline` for the new single-level model and compare it to the old single-level model (from 1).i). Which model's residuals ($\epsilon$) fulfil the assumptions of the General Linear Model better?)
    iv. Also make a quantile-quantile plot for the residuals of the  multilevel model with two intercepts. Does it look alright?
    
```{r}
#New dataframe and also averaging f0mn
df2 <- politeness %>% 
  filter(!is.na(f0mn)) %>% 
  group_by(subject, gender) %>% 
  summarise(mean_f0mn = mean(f0mn))

#We had NAs because some participants have blank rows. Debate over whether we should remove the NA rows and average over the existing rows, or completely remove the participants? We ended up just omitting the NA rows and keeping the participants, just averaging from the existing datapoints they had

#Making a model from the new dataframe
df2m <- lm(mean_f0mn ~ gender, data = df2)

qqnorm(residuals(df2m))
qqline(residuals(df2m))

#Comparing to m1
qqnorm(residuals(m1))
qqline(residuals(m1))

#Very difficult to determine visually. m1 has some weird outliers and is a bit curvy, but has many more datapoints that are completely on the line. df2m has much fewer data points that aren't exactly on the line, but pretty close
#Buuut perhaps m1 is a little bit better, because it has sooo many points right on the line

#Making a qq-plot for m4 (the multilevel model with two intercepts)
qqnorm(residuals(m4))
qqline(residuals(m4))
#It looks very similar to m1, and looks okay ish
```

3) Plotting the two-intercepts model
    i. Create a plot for each subject, (similar to part 3 in Exercise 1), this time also indicating the fitted value for each of the subjects for each for the scenarios (hint use `fixef` to get the "grand effects" for each gender and `ranef` to get the subject- and scenario-specific effects)
    
```{r}
ggplot(politeness, aes(scenario, f0mn, colour = attitude)) +
  geom_point() +
  geom_point(aes(fix, ran$subject), shape = "diamond") +
  geom_point() +
  facet_wrap(~subject)

#WHY IS RAN A LIST AND WHY CANT I CHANGE IT?

#Mina's answers:
fitted <- fitted(m4)
politeness_na_removed <- politeness %>% 
  na.omit()
politeness_na_removed$fitted_f0mn <- fitted
ggplot(politeness_na_removed, (aes(x = scenario, y = f0mn, color = attitude)))+ 
  geom_point()+
  geom_point(aes(scenario, fitted_f0mn), color = "darkgrey", shape = 17)+
  facet_wrap(.~subject)+ 
  theme_bw()
```
    
    
## Exercise 3 - now with attitude

1) Carry on with the model with the two unique intercepts fitted (_scenario_ and _subject_).
    i. now build a model that has _attitude_ as a main effect besides _gender_
    ii. make a separate model that besides the main effects of _attitude_ and _gender_ also include their interaction
    iii. describe what the interaction term in the model says about Korean men's pitch when they are polite relative to Korean women's pitch when they are polite (you don't have to judge whether it is interesting)  

```{r}
#Adding attitude as a main effect
m5 <- lmer(f0mn ~ gender + attitude + (1 | scenario) + (1 | subject), data = politeness)
#Adding the interaction of gender and attitude
m6 <- lmer(f0mn ~ gender * attitude + (1 | scenario) + (1 | subject), data = politeness)
```

2) Compare the three models (1. gender as a main effect; 2. gender and attitude as main effects; 3. gender and attitude as main effects and the interaction between them. For all three models model unique intercepts for _subject_ and _scenario_) using residual variance, residual standard deviation and AIC.  

```{r}
#Is R^2 the residual variance?
r.squaredGLMM(m5)
r.squaredGLMM(m6)

#Residual standard deviation
sigma(m5)
sigma(m6)

#AIC
AIC1 <- AIC(logLik(m5))
AIC2 <- AIC(logLik(m6))
```

3)  Choose the model that you think describe the data the best - and write a short report on the main findings based on this model. At least include the following:
  i. describe what the dataset consists of  
  ii. what can you conclude about the effect of gender and attitude on pitch (if anything)?  
  iii. motivate why you would include separate intercepts for subjects and scenarios (if you think they should be included)  
  iv. describe the variance components of the second level (if any)  
  v. include a Quantile-Quantile plot of your chosen model  